%\clearpage

\addcontentsline{toc}{chapter}{Acknowledgements}

\begin{acknowledgements}
    Alhamdulillah, I'd like to express my profound gratitude to Almighty \textbf{Allah}, for His bounties and blessings to complete this dissertation. Then, I'd like to convey my heartiest thanks to all who were with me during this journey towards achieving my doctoral degree at RWTH Aachen University, Germany. Foremost, I'd like to acknowledge my sincere gratitude to my supervisor Prof. Dr. \textbf{Stefan Decker} for hiring me as a Scientist at Fraunhofer FIT and supervising my Doctoral thesis at the RWTH Aachen University. His insightful comments and encouraging remarks addressed during my each presentation patience and supportive guidance throughout my graduate studies at the university. His kind motivation and profound direction allowed me to explore and learn while maintaining a steady course towards the destination. He introduced me to the research areas and provided me with the proper understanding with his expertise. His suggestions and invaluable advice gave subtle direction to my efforts. In a nutshell, I found him very friendly, cooperative, very open-minded, and approachable. 
    %Believe me, I have never been depressed while I was working at Fraunhofer FIT, Germany. 
    
    \hspace*{5mm} Meanwhile, my heartfelt gratitude and thanks endlessly go to my advisors: Dr. Oya Beyan and Dr. Michael Cochez. The door to their offices was always open for me whenever I ran into any difficulties. They consistently motivated me through their domain expert, guidance, and insights and steered me in the right direction whenever I needed it. To Dr. \textbf{Oya Beyan}, I'm very very grateful to her, and I express my sincere gratitude to her for hiring me as a PhD student in Ireland as well as for making me aware of my current research position at Fraunhofer FIT. %I'm forever indebted to her, as I consider her as my key to Europe. 
    To Dr. \textbf{Michael Cochez}, I found him one of my best research mentors, who always thought and motivated me how to research in a structured way, how to think scientifically, and how to write scientific articles. He was not only a mentor to me, but also a good human being whom I could share my research interests and personal matters without hesitations. To the examiner, Prof. Dr. \textbf{Dietrich Rebholz-Schuhmann}, I'm very grateful and indebted to him too for being one of my best mentors and motivating leaders. He has always inspired me since the beginning of my journey in Ireland as well as in Germany. 
    
    \hspace*{5mm} Overall, I am gratefully indebted to all of the above for their prompt, thorough and timely feedback, corrections, and valuable suggestions on this thesis. To be honest, without their passionate participation and immense knowledge in the research and science, solving complex could not have been impossible. Next, I am incredibly grateful indebted to my parents \textbf{Abdur Razzaque} and \textbf{Monoara Begum} for their sacrifice and care throughout my life, especially to my father who had a dream that one day I'll finish my Ph.D., which is why I dedicate my thesis to my father. Nevertheless, I am very grateful to my loving wife \textbf{Saroar Zahan} and my son \textbf{Ahmed Shadman} for their patience, contributions, and sincere accompany here in Germany. Their unconditional love and never-ending support gave me the confidence to utilize my abilities. Their prayer and guidance is the reason for where I stand here today. Besides, I would like to thank all of my family members, including my sister Josna Khatun and my brother Mamtaz Uddin, whose immense love, care, and encouragement always gather inspiration in my work and silent acceptance of my long absence from my home country. 
    
    \hspace*{5mm} I would like to graciously acknowledge the help of my current members of Data Science \& AI Department at Fraunhofer FIT and Chair of Computer Science 5 - Information Systems and Databases at RWTH Aachen University, for providing friendly environment at the working place. First, I would like to express my sincere indebtedness to the head of the department Dr. \textbf{Christoph Lange} and other colleagues, including Till D{\"o}hmen, J{\'o}ao Bosco Jares, Yeliz Ucer, Lars Gleim, and Sascha Welten for their expertise and patient guidance throughout this research work. To \textbf{Till D{\"o}hmen}, I'm very thankful for translating me the English abstract of this thesis into German. 
    
    \hspace*{5mm} Besides, I extend my sincere thanks to IT help-desk of Fraunhofer FIT for providing me access to it's computing infrastructure in which I performed most of the experiment, including data processing, storing, training, and analysis. Additionally, I am grateful for the support and help from Mr. Reinhard Linde, Ms. Leany Maassen, Ms. Romina Reddig, and Ms. Tatiana Liberzon. Mr. Reinhard Linde was always accessible and helped create network and computing infrastructures, which substantially boosted the progress of my works. Furthermore, I am grateful to have access to PanCancerAtlas project from The Cancer Genome Atlas~(TCGA). Nevertheless, writing this thesis is made more accessible by incredible efforts by many open source communities and documentations provided by TensorFlow, Keras, Scikit-learn, Python, Apache Spark, RuleMatrix, Skater, ExplainX, and SHAP community and all those who have contributed to APIs, whose work ultimately brought the machine learning to the masses! 
    
    \hspace*{5mm} Nevertheless, I want to acknowledge the help and care of every member of the Bangladeshi Community in Germany. I can never forget the pleasant moments and memories that I passed with this community during my stay here in Germany, including Mr. Sami Chowdhury and Mr. Mesbah Saber, for their love, support and inspiration in removing my loneliness in this place by providing a family-like environment. Finally, I'd like to graciously acknowledge the help and encouragement that I obtained from my childhood friends and well-wishers, including Naimul, Khairul, Rakib, and Mahidul throughout this research work. \\ \\ %I would like to express my heartfelt thanks to all of them for being with me with immense support and care and to make this work successful.
    \\ \\ \\ 
    \flushright Md. Rezaul Karim, MSc. \\
    Computer Science 5 - Information Systems and Databases\\
    RWTH Aachen University\\ 
    Aachen, Germany \\
    March 2021
\end{acknowledgements}